Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features

### 2016-12-16 paper/structureScene.tex
New conclusion

In this paper we consider the benefit of 1) the scattering transform for building discriminant features and 2) unsupervised clustering for the modeling of large-scale environmental acoustic scenes.
%This leads to improvements in both the ASSR and ASC tasks.

Representing a scene using a reduced set of representatives, with adapted similarity strategies, allows us to improve ASSR performance compared to the early-integration scheme which have been demonstrated to be equivalent to BOF approaches using GMMs on well-balanced datasets \cite{lagrange:hal-01082501}. These results advocate for a hierarchical model of the scene, where short-time structure--\ie below a second--is represented using a convolutional network such as the scattering transform, whereas larger time scales--\ie over a second--are modeled as adaptive aggregates.

Within a supervised setting of acoustic scene classification (ASC), we have shown the comparative benefits of log-scattering features over MFCCs, of majority voting at a mid-level time scale of $T=372\,\mathrm{ms}$ over prediction at a high-level time scale of $T=24\,\mathrm{s}$, and of an adapted Gaussian kernel over the linear kernel in support vector machine classification for the tasks we considered.
%, identified in the time domain in the unsupervised setting and in the features domain in the supervised case, provided that the dimensionality is sufficiently high. \ja{I'm not sure what this last part means.}

These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in ecoacoustics.

>>>>

This paper presents a new approach for modeling acoustic scenes which utilizes scattering transforms at small scales and a cluster-based representation at large scales. Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature. We demonstrate the usefulness of scattering-based representations for auditory scenes and the value of late integration using experiments in auditory scene classification. In order to extend the ideas of late integration to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval. For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in ecoacoustics.
